---
title: "Determine Zooniverse Leaderboard"
output: html_notebook
---

We need a set of scripts that we can use to look at a zooniverse classification file and determine who has made the most classifications in a particular time span. This code will do jus that.

Set things up
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(lubridate)
library(rstudioapi)

myFunctions<-list.files(here("Code", "Functions"), pattern = "R") 
#points to files that need to be sourced

sapply(here("Code", "Functions", myFunctions), source) #sources all of the functions in the Functions folder.
```
### Step 1. Choose the classification files to work with.
We have multiple classification files. The `choose_my_classification_file` function asks the user for input so that they can select the classification file to work with. The classification file should be on the t://drive in Bart_s04/Research_group/NoCoWild/Data/classification_files.



```{r}
animals<-choose_file_isolate_workflow()
species<-choose_file_isolate_workflow()
```
Now isolate each file to the date ranges of interest using `narrow_to_date.R`

```{r}
animals<-narrow_to_date(animals[[1]])
species<-narrow_to_date(species[[1]])
```
### Step 2. Now, for each data frame, identify users and then quantify classifications per user per time.

#### Start with Is there an animal workflow ("animals"):
Identify users

How many are there?
```{r}
unique(animals$user_name)
length(unique(animals$user_name))
```
Now combine all the not-logged-in users to a single username.

```{r}
pattern<-"not-logged-in"
replacement<-grep("not-logged-in",animals$user_name)
#identifies all rows with user name including not logged
#in

#now reassign factor name
animals$user_name[replacement]<- "Not logged in"

unique(animals$user_name)
```
Now assign each day to a week of the year

```{r}
animals<-animals %>% mutate(
  week=isoweek(created_at)
)
```
Now summarize effort by user by week
```{r}
a_Effort<-animals %>% group_by(cYear, week, user_name) %>% summarize(
  Num_Class = length(classification_id),
  max_date = max(created_at)
)

a_Effort2022<-Effort %>% filter(cYear == 2022)
```

#### Now do similar for species identification

Continue with Species identification workflow ("species"):

How many are there?
```{r}
unique(species$user_name)
length(unique(species$user_name))
```
Now combine all the not-logged-in users to a single username.

```{r}
pattern<-"not-logged-in"
replacement<-grep("not-logged-in",species$user_name)
#identifies all rows with user name including not logged
#in

#now reassign factor name
species$user_name[replacement]<- "Not logged in"

unique(species$user_name)
```
Now assign each day to a week of the year

```{r}
species<-species %>% mutate(
  week=isoweek(created_at)
)
```
Now summarize effort by user by week
```{r}
s_Effort<-species %>% group_by(cYear, week, user_name) %>% summarize(
  Num_Class = length(classification_id),
  max_date = max(created_at)
)

s_Effort2022<-s_Effort %>% filter(cYear == 2022)
```

