---
title: "Image Pre-processing 1: Assign images to events"
output: html_notebook
---
## Introduction
This is the R code needed to use the metadata from a folder of images to compare the image time stamps and group images into their respective "events."

An event is a set of pictures taken within a time threshold.  We set our cameras to take up to 3 pictures in a bout, and set them with a 1 minute delay before they can take more pictures, so we can use 1 minute as our threshold.

### Before you begin

In order to run this code, you will need to have the following in place:

* A folder with a set of images (and only images) in it.
* A csv file made using *exiftool* that has date/time metadata from all of the photos in the folder


### Step 1
Set up the workspace
```{r, results='hide'}
rm(list = ls())
library(tidyverse)
library(lubridate)
library(here) #helps keep directory structure clean when using .Rmd files
```
### Step 2
First set up the path to the folder containing the images and the folder containing the metadata file.

```{r}
exif_path<-"/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/exif_metadata/"
image_path<-"~/Dropbox/Nature Up North/Citizen Science/North Country Wild/code_testing/Processed"
```

Now create a variable that is the list of image file names. 

```{r}
files<-list.files(image_path) #creates variable called files as vector of filenames
```
### Step 3 
Now we need to pull in the image metadata.  It was generated and saved as a .csv outside of R using a tool called *exiftool* that is run from the command line. 

```{r}
exif_file_name<-"exif_test.csv" #put correct name of file here.
Exif <- read.csv(paste0(exif_path,exif_file_name)) #obviously, put correct file name here

#may throw an error if your connection to the t:drive drops.
```

### Step 4
Now that we have the metadata, we can see that the DateTimeOriginal column came in as a character.  Let's change it with lubridate. **A very important step is to ORDER the data frame in terms of date/time.**  If you fail to order the data frame, when you compare dates/times you will get incorrect output.

```{r}
Exif$DateTimeOriginal<-ymd_hms(Exif$DateTimeOriginal)
Exif<-Exif[order(Exif$DateTimeOriginal),]#THIS IS KEY
```
Now see if we can calculate the time differences among the photos

```{r}
MyDiffs<-diff(Exif$DateTimeOriginal)
#returns what looks like an empty variable in the environment window
MyDiffs<-as.numeric(MyDiffs) 
#now it shows up as numeric data representing the number of seconds between the image and the next image (in temporal order).Notice there is one less value in MyDiffs than there are images in the folder.
```
### Step 5 
Now assign images to events using this logic.  We will need to set the time threshold (default is 1 minute).

```{r}
thresh<-60 #sets a threshold for number of seconds between events
ev<-vector() #this will store event numbers temporarily
L <- length(files) #determines number of image files in folder
Event_num<-numeric()
#now use a loop to populate the ev vector with diffs..
interval<-diff(Exif$DateTimeOriginal)
ev<-1
cond<- interval > thresh #sets condition for assigning to a new event where minutes corresponds to thresh

#now write the loop that will assign each image to the proper event.

   for(i in 1:(L-1)){
      if(!cond[i]) ev<-ev else ev<-ev+1
      Event_num<-c(Event_num,ev)
   }
  Event<-c(1,Event_num)
```
Now append the Event to the data frame csv to capture the event assignment for each image.
```{R}
Exif$Event<-Event
```
### Step 6
Now export Exif for use in assigning images to "image numbers" in the zooniverse manifest style.  Use a file naming convention based on the name of the original exif data so that it is clear what set of images this file is for.

Strip off the "exif_" lead to the file name and replace with "images_by_event_"
```{r}
exif_file_name<-str_replace(exif_file_name, "exif_", "images_by_event_") #finds "exif_" in file name, replaces it with "images_by_event.
```
Now we need to save the file. This should go into the project Output -> PreZooniverse -> ImagesByEvent folder.

```{r}
output<-here("Output","PreZooniverse","ImagesByEvent")  
write.csv(Exif, paste0(output, "/",exif_file_name))
```



  
