---
title: "Consensus picks"
output: html_notebook
---

Once we have flattened all of our data, we need to examine it to determine the consensus species ID for each event.

## Set up the workspace and read in the data
```{r}
rm(list = ls())
library(here)
library(tidyverse)
path_to_flat<-here("Output", "PostZooniverse", "Flattened")
flat_files<-list.files(path_to_flat, pattern =".csv")

DF <- read.csv(paste0(path_to_flat, "/",flat_files), stringsAsFactors=F)
```

```{r}
#get rid of data from beta test
DF<-DF %>% filter(round !="beta test")
```

Summarize the data to see the number of distinct subjects (events) and the number of distinct classifications in the data set. Then create a new column that has a count of how many unique classifications were made.
```{r}
DF %>% summarise(n_distinct(subject_ids), n_distinct(classification_id)) 


DF<- DF %>% 
     group_by(subject_ids) %>% # count up the number of distinct classification IDs
     mutate(., 
            num_class = n_distinct(classification_id),
            num_species = n_distinct(choice)) %>% #because there will be >1 row per classification_id if >1 spp
     arrange(., subject_ids, classification_id) 
DF %>% View


```
Look at those cases for which the number of species was 5
```{r}
DF %>% filter(num_species > 4)
```

```{r} 
DF %>% arrange(desc(num_species, subject_ids)) %>% View()
```

Now that we know the number of species identified per subject id, we need a way to step through and determine the consensus species for each subject id. First we'll write some code to check the species counts per subject id and classification id.
## Subject level metrics
```{r}
subject_metrics <- DF %>% ungroup %>%
     group_by(., subject_ids) %>%
     mutate(., num_votes = n(), # if a  user ids >1 spp, there will be more votes than classifications
            diff_species = n_distinct(choice)) # count the total number of different species reported by different users, for pielous score

```

```{r}
species_counts <- DF %>% ungroup %>%
     group_by(subject_ids, classification_id) %>%
     summarise(total_spp_by_user = mean(num_species)) %>% #Need to select only one row per classification_id, then summarise across those. 
     summarise(., agg_num_species = round(median(total_spp_by_user), 0))#aggregate species count, which is median rounded up
     glimpse(species_counts)

cleaned_classifications <- left_join(subject_metrics, species_counts) %>% ungroup
glimpse(cleaned_classifications)
```
The group of subject ids for which there was > 1 species assigned is evident by filtering species_counts to extract all of those subject id's with > 1 assigned species.  Call this value "checkSubjects".

```{r}
checkSubjects <- filter(species_counts, agg_num_species > 1)
```
There are 442 subject id's with > 1 species assigned.

## Species level metrics
```{r}
grouped_classifications <- cleaned_classifications %>% 
     select(., -num_species) %>% # these aren't relevant
     group_by(., subject_ids, num_class, num_votes, agg_num_species, diff_species, choice) # fields at subject level or higher
```


#Tally the votes for each species ID'd within a subject
This yields species_votes, which tells us, for each subject_id, the number of classifications and the proportion of the vote received by each species for each subject_id. We can use this to run a list of subjects for which we need to go back and confirm the species.
```{r}
species_votes <- grouped_classifications %>% 
     # for every species within a subject, aggregate votes.
     summarise(., votes = n_distinct(classification_id)) %>% #count up the number of votes per species choice
     mutate(propvote = votes/sum(votes), #calculate proportion of votes for this species
            propclass = votes/num_class) #calculate proportion of classifications for this species
```
When num_votes > num_class you have a situation where there was more than one species detected in an event (typically Gus and one or more camera trappers.). In these cases, propvote will not equal propclass, and the sum of propclass for that subject id will be greater than 1.0 (though the sum of propvote will be equal to 1.)

Let's start by filtering species_votes to extract just the subjects mentioned in checkSubjects.

```{r}
Subjects_to_check<-checkSubjects$subject_ids

species_votes_to_check<- species_votes %>% ungroup() %>%  filter(subject_ids %in% Subjects_to_check) #need %in% because i'm filtering on a vector rather than a condition or single value.
```
Now we want to filter out the cases for which propvote = propclass and propclass is 0.8 or greater or 0.2 or smaller.

```{r}
species_votes_to_check<- species_votes_to_check %>% filter((propvote == propclass) & (between(propclass, 0.21, 0.79)))
```
Now get the list of the subject ids in species_votes_to_check and use it to filter cleaned_classifications to get the list of subjects, images, events etc. to check.

```{r}
Subjects_to_filter<-unique(species_votes_to_check$subject_ids)
```
And now filter cleaned_classifications by those data

```{r}
filesToCheck<-filter(cleaned_classifications, subject_ids %in% Subjects_to_filter)
```
And now summarize
```{r}
filesToCheck<-filesToCheck %>% group_by(subject_ids) %>% summarize(
  Event = Event,
  Img1 = Imj1,
  Img2 = Imj2,
  Img3 = Img3,
  round = round,
  CamNum = CamNum,
  SD_card = SD_card_num
)
```
Now loop through filesToCheck to grab just the first instance of each subject id and associated variables.  

Make data frame to put results in 
```{r}
Subjects<-unique(filesToCheck$subject_ids)

Checklist<-data.frame(subject_ids = integer(255),
                      Event = integer(255), 
                      Img1 = character(255), 
                      Img2 = character(255), 
                      Img3 = character(255),
                      round = character (255),
                      CamNum = integer(255), 
                      SD_card = integer(255))
```
Now loop
```{r}
for(i in 1:length(Subjects)){
  mySample<-filter(filesToCheck, subject_ids == Subjects[i])
  Checklist$subject_ids[i]<-Subjects[i]
  Checklist$Event[i]<-unique(mySample$Event)
  Checklist$Img1[i]<-unique(mySample$Img1)
  Checklist$Img2[i]<-unique(mySample$Img2)
  Checklist$Img3[i]<-unique(mySample$Img3)
  Checklist$round[i]<-unique(mySample$round)
  Checklist$CamNum[i]<-unique(mySample$CamNum)
  Checklist$SD_card[i]<-unique(mySample$SD_card)
}
```
Now save

```{r}
write.csv(Checklist, "files_to_check_no_beta_test.csv")
```

Now we need to assign the correct species for the files that we are not checking:
Extract from cleaned classifications all of the subject ids that are NOT part of the list that needs to be checked:

```{r}
Subjects_to_keep<-species_votes %>% filter(propvote >= 0.8)
Subjects_to_keep <-Subjects_to_keep$subject_ids


NeedAssignment<-cleaned_classifications %>% filter(subject_ids %in% Subjects_to_keep)
```
Now create data frame to accept final assignments

```{r}
Assigned<-data.frame(subject_ids = integer(1421),
                      Event = integer(1421),
                     Img1 = character(1421), 
                      Img2 = character(1421), 
                      Img3 = character(1421),
                      round = character (1421),
                      CamNum = integer(1421), 
                      SD_card = integer(1421),
                     choice = character (1421))
```
Now loop
```{r}
NeedAssignment$choice<-as.character(NeedAssignment$choice)
for(i in 1:length(Subjects_to_keep)){
  mySample<-filter(NeedAssignment, subject_ids == Subjects_to_keep[i])

  Assigned$subject_ids[i]<-Subjects_to_keep[i]
  Assigned$Event[i]<-unique(mySample$Event)
  Assigned$Img1[i]<-unique(mySample$Imj1)
  Assigned$Img2[i]<-unique(mySample$Imj2)
   
  Assigned$Img3[i]<-unique(mySample$Img3)
  
  Assigned$round[i]<-unique(mySample$round)
 
  Assigned$CamNum[i]<-unique(mySample$CamNum)
  
  Assigned$SD_card[i]<-unique(mySample$SD_card_num)
  
  Assigned$choice[i]<-unique(mySample$choice[1])
 
}
```
Now save

```{r}
write.csv(Assigned, "AssignedSpecies_no_beta_test.csv")
```








