---
title: "Image Pre-processing 1: Assign images to events"
output: html_notebook
---
## Introduction
This is the R code needed to use the metadata from a folder of images to compare the image time stamps and group images into their respective "events."

An event is a set of pictures taken within a time threshold.  We set our cameras to take up to 3 pictures in a bout, and set them with a 1 minute delay before they can take more pictures, so we can use 1 minute as our threshold.

### Before you begin

In order to run this code, you will need to have the following in place:

* A folder with a set of images (and only images) in it.
* A csv file made using *exiftool* that has date/time metadata from all of the photos in the folder


### Step 1
Set up the workspace
```{r, results='hide'}
rm(list = ls())
library(tidyverse)
library(lubridate)
library(here) #helps keep directory structure clean when using .Rmd files
myFunctions<-list.files(here("Code", "Functions")) 
#points to files that need to be sourced

sapply(here("Code", "Functions", myFunctions), source) #sources all of the functions in the Functions folder.

output<-here("Output","PreZooniverse","ImagesByEvent") #sets path for file output.

```
### Step 2
First set up the path to the folder containing the images and the folder containing the metadata file.

```{r}
exif_path<-"/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/exif_metadata/"
image_path<-"~/Dropbox/Nature Up North/Citizen Science/North Country Wild/code_testing/Processed"
```

Now create a variable that is the list of image file names. 

```{r}
files<-list.files(image_path) #creates variable called files as vector of filenames
```
### Step 3 
Now we need to pull in the image metadata.  It was generated and saved as a .csv outside of R using a tool called *exiftool* that is run from the command line. 

```{r}
exif_file_name<-"exif_2021_R1_C024_SD100_uneven.csv" #put correct name of file here.
Exif <- read.csv(paste0(exif_path,exif_file_name)) #obviously, put correct file name here

#may throw an error if your connection to the t:drive drops.
```

### Step 4
Now assign images to events using assignEvents function (assignEvents.R).  We will need to set the time threshold (default is 1 minute).

```{r}
Exif<-assignEvents(Exif)
```

### Step 5
Now export Exif for use in assigning images to "image numbers" in the zooniverse manifest style.  Use a file naming convention based on the name of the original exif data so that it is clear what set of images this file is for.

Strip off the "exif_" lead to the file name and replace with "images_by_event_"
```{r}
exif_file_name<-str_replace(exif_file_name, "exif_", "images_by_event_") #finds "exif_" in file name, replaces it with "images_by_event.
```
Now we need to save the file. This should go into the project Output -> PreZooniverse -> ImagesByEvent folder. I set that up by defining `output` in the file setup step at the top of the page.

```{r}
write.csv(Exif, paste0(output, "/",exif_file_name))
```
At the moment all of this code works!


  
