---
title: "Flattening our Zooniverse files"
output: html_notebook
---
The code in this file is designed to open the subjects export file and pull the columns that contain the subject_id for each image as well as the subject_set_id so that we can keep add that information to the classification data.

Set things up:
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(rstudioapi)

myFunctions<-list.files(here("Code", "Functions"), pattern = "R") 
#points to files that need to be sourced

sapply(here("Code", "Functions", myFunctions), source) #sources all of the functions in the Functions folder.
```

Now pull in data file:
```{r}

rstudioapi::showDialog(
    title = "Let's give R some file information",
    message = "Select the file containing the subject export data you wish to process")

  #path to exif file folder on t:drive
  path_to_class_file<-"/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/classification_files/"

  class_file_path<-rstudioapi::selectFile(
    path = "/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/classification_files/",
    filter = "CSV files (*.csv)",
    label= "Select desired file"
  )

  DF<-read.csv(class_file_path)
  filename<-basename(class_file_path)

```
Now filter to the workflow id that you want
```{r}
unique(DF$workflow_id)
```
```{r}
newDF<-filter(DF, workflow_id == 16695)
length(unique(newDF$subject_id))
```



Now we just need to extract the two columns subject_id and subject_set_id into a new df and write it.

```{r}
newDF<-DF %>% select(subject_id, subject_set_id)
```
Now save
```{r}
write.csv(newDF, here("Data", "subject_set_ids-2023-06-13.csv"))
```

