---
title: "check-z-subject-sets"
format: html
editor: visual
---

## Check zooniverse subject set progress

You may encounter a situation in which you need to prepare a subject set or subject sets to upload to the Zooniverse, but you are curious about the status of existing subject sets (have they been fully classified?).

To check, go to the "behind the scenes" region of our zooniverse project by logging in and then choosing the "Build a project" menu.

Go to Data Exports on the left-hand side and request a new subject export.

Once you have the subject export, rename the downloaded file with the date you exported the subjects, e.g. 2025-12-16-subjects.csv.

Place the file in Data –\> Post Zooniverse –\> Subjects.

You can then use the following code to evaluate the subject set.

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(rstudioapi)
```

Now navigate to the file you want to consider

```{r}
path <- rstudioapi::selectFile(
  caption = "Select subjects file", 
  filter = "CSV Files (.csv)")
subjects <-read_csv(path)
```

Now get names of zooniverse subject sets to join

```{r}
set_names <- read_csv(here("Data/PostZooniverse/Subjects/zooniverse_subject_set_ids.csv"))
```

Now get some summary information

```{r}
subject_summary <- subjects |> group_by(subject_set_id, workflow_id) |>
  summarise(
    num_subjects = length(unique(subject_id)),
    #returns correct number of images in subject set
    num_retired = sum(!is.na(retired_at))
  ) 
View(subject_summary)
```

Now join

```{r}
subject_summary_join <- 
  left_join(
    subject_summary, 
    set_names |> select(subject_set_id, original_zooniverse_name),     
    by = "subject_set_id")|>
  relocate(original_zooniverse_name, .after = subject_set_id)
```

Then save this file
```{r}
write.csv(subject_summary_join, here(paste0("Output/subject-set-reports/",Sys.Date(),"-subject-set-report.csv")), row.names = F)
```

