---
title: "Comparing datasets"
output: html_notebook
---

I'm curious to look at the range of dates for each of the 3 data sets I just made so I can see if I can decipher which projects are split between classification files.  

Begin by cleaning things up and pulling in data

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```
Pull in the data files

```{r}
DS<-read.csv(here("Output", "PostZooniverse", "Flattened", "north-country-wild-classifications.6June2022_flat_2022-06-09_wf1_donovan.csv"))

KA<-read.csv(here("Output", "PostZooniverse", "Flattened", "north-country-wild-classifications.6June2022_flat_2022-06-09_wf1.1_kate.csv"))

CWMM<-read.csv(here("Output", "PostZooniverse", "Flattened", "north-country-wild-classifications.6June2022_flat_2022-06-09_sp_id_cole_maggie.csv"))
```

Now get rid of the answers columns that follow choice
```{r}
DS<-DS %>% select(!starts_with("answers"))
KA<-KA %>% select(!starts_with("answers"))
CWMM<-CWMM %>% select(!starts_with("answers"))

DS<-DS %>% select(!starts_with("task"))
KA<-KA %>% select(!starts_with("task"))
CWMM<-CWMM %>% select(!starts_with("task"))
```
Check that all the column names are the same in each data frame

```{r}
colnames(CWMM) == colnames(DS)
colnames(DS) == colnames (KA)
```
Now add a column to each dataframe indicating its source (i.e the dataframe)
```{r}
DS$Source<-"DS"
KA$Source<-"KA"
CWMM$Source<-"CWMM"
```
Now get rid of the beta test classifications from CWMM

```{r}
beta<-which(CWMM$round1 == "beta test")
#now delete those rows
CWMM<-CWMM[-beta,]
```
Now rearrange some of the data so that we can get rid of redundant columns

Duplicate types of columns = 
round and round1
Img1, Img2, Img3 and Img4, Img5, Img6
CamNumber, CamNum, CamNum2
SDNumber, SD_card_num, SDCard
Forest and Forest1
Event and Event1

Need to check if these have data in each data frame
```{r}
length(unique(DS$Img1))
length(unique(KA$Img1))
length(unique(CWMM$Img1))

length(unique(DS$Img4))
length(unique(KA$Img4))
length(unique(CWMM$Img4))
```
DS and KA have no images in Img4, Img5, Img6.
CWMM has images in both Img1 - Img3 and Img4-Img6 this is cole's data

```{r}
Cole<-which(!is.na(CWMM$Img4))
is.na(CWMM[Cole,]$Img1)
CWMM[Cole,]$Img1<-CWMM[Cole,]$Img4
CWMM[Cole,]$Img2<-CWMM[Cole,]$Img5
CWMM[Cole,]$Img3<-CWMM[Cole,]$Img6
Maggie<-which(is.na(CWMM$Img4))
length(!is.na(CWMM[Maggie,]$Img1))
length(!is.na(CWMM$Img1))
```
Ok, successfully  moved all images into Img1, Img2, Img3 columns so can delete Img4, Img5, Img6 from each data frame.

```{r}
CWMM<-CWMM %>% select(-c(Img4:Img6))
DS<-DS %>% select(-c(Img4:Img6))
KA<-KA %>% select(-c(Img4:Img6))
```
Now deal with round and round1
```{r}
which((is.na(DS$round))==TRUE)
length(which((is.na(DS$round1))== TRUE))
```
DS has no values in round1
```{r}
which((is.na(KA$round))==TRUE)
length(which((is.na(KA$round1))== TRUE))

```
KA has no values in round1

Now check CWMM
```{r}
length(which((is.na(CWMM$round))))
```
All but one row of CWMM lack values in round
Which row has a value?
```{r}
View(CWMM[(which(!is.na(CWMM$round))),])
```
One value from the older data set row 8368
So, except for that row, move all of round1 into round in CWMM. Or, just reassign round = 1 to that row after moving. That would be easier.

```{r}
CWMM$round<-CWMM$round1
CWMM$round[8368]<-1
```

Now I can get rid of round1 in all data sets
```{r}
CWMM<-CWMM %>% select(-round1)
DS<-DS %>% select(-round1)
KA<-KA %>% select(-round1)
```

Now deal with CamNumber, CamNum, CamNum2
```{r}
length(which(is.na(DS$CamNum)))
length(which(is.na(DS$CamNum2)))

length(which(is.na(KA$CamNum)))
length(which(is.na(KA$CamNum2)))
```
Neither DS nor KA have values in CamNum or CamNum2

What about CWMM
```{r}
length(which(is.na(CWMM$CamNumber)))
length(which(is.na(CWMM$CamNum)))
length(which(is.na(CWMM$CamNum2)))

```
Except for that one row (likely 8368) in CWMM, there are no data in CamNumber, but are data in both CamNum and CamNum2 that need to be moved.

```{r}
CamNum<-which(!is.na(CWMM$CamNum))
```
Looks like Cole's data has values in CamNum
Move to CamNumber
```{r}
CWMM$CamNumber[CamNum]<-CWMM$CamNum[CamNum]
```
Now fix CamNum2
```{r}
CamNum2<-which(!is.na(CWMM$CamNum2))
CWMM$CamNumber[CamNum2]<-CWMM$CamNum2[CamNum2]
```

Now I can get rid of CamNum and CamNum2
```{r}
CWMM<-CWMM %>% select(-c(CamNum, CamNum2))
DS<-DS %>% select(-c(CamNum, CamNum2))
KA<-KA %>% select(-c(CamNum, CamNum2))
```
Now deal with SD card
```{r}
length(which(is.na(DS$SDNumber)))
length(which(is.na(DS$SD_card_num)))
length(which(is.na(DS$SDCard)))
```
DS has no values in SD_card_num or SDCard
```{r}
length(which(is.na(KA$SDNumber)))
length(which(is.na(KA$SD_card_num)))
length(which(is.na(KA$SDCard)))
```
KA is missing SD card numbers for about half the observations in SDNumber but has no values in SD_card_num or SDCard so those columns can be deleted.

What about CWMM?
```{r}
length(which(is.na(CWMM$SDNumber)))
length(which(is.na(CWMM$SD_card_num)))
length(which(is.na(CWMM$SDCard)))
```
CWMM has no values in SD card. That column can be deleted from all three data sets.Maggie's data appear not to have values in SD_card_num. So basically I need to move the values of SD_card_num into SDNumber
```{r}
SDfix<-which((!is.na(CWMM$SD_card_num)))
CWMM$SDNumber[SDfix]<-CWMM$SD_card_num[SDfix]
```
Ok. Now delete extra SD columns

```{r}
CWMM<-CWMM %>% select(-c(SD_card_num, SDCard))
DS<-DS %>% select(-c(SD_card_num, SDCard))
KA<-KA %>% select(-c(SD_card_num, SDCard))
```

At this point, it might be worth adding a student to each of the data sets. I'll need to partition CWMM between Cole and Maggie.

```{r}
DS$Student<-"Donovan"
KA$Student<-"Kate"
CWMM$Student<-NA
CWMM$Student[Cole]<-"Cole"
CWMM$Student[Maggie]<-"Maggie"
which(is.na(CWMM$Student))
```
Ok, now deal with forest and forest1
```{r}
length(which(is.na(DS$Forest)))
length(which(is.na(DS$Forest1)))

length(which(is.na(KA$Forest)))
length(which(is.na(KA$Forest1)))

```
Neither DS nor KA have values in Forest1. Can be deleted from those two.
Check CWMM

```{r}
length(which(is.na(CWMM$Forest)))
length(which(is.na(CWMM$Forest1)))
```
Maggie's data, but not Cole's, have info in Forest1

Fix that here
```{r}
maggie_trees<-which(!is.na(CWMM$Forest1))
CWMM$Forest[maggie_trees]<-CWMM$Forest1[maggie_trees]
```
Now get rid of forest 1 from all
```{r}
DS<-DS %>% select(-Forest1)
KA<-KA %>% select(-Forest1)
CWMM<-CWMM %>% select(-Forest1)
```

Now deal with Event and Event1 by just deleting them as I don't need them.
```{r}
DS<-DS %>% select(-c(Event,Event1))
KA<-KA %>% select(-c(Event,Event1))
CWMM<-CWMM %>% select(-c(Event,Event1))
```
From 48 variables down to 32! 

I can also delete Num, Lat, Long
```{r}
DS<-DS %>% select(-c(Num, Lat, Long))
KA<-KA %>% select(-c(Num, Lat, Long))
CWMM<-CWMM %>% select(-c(Num, Lat, Long))
```
Now I can save each data set and also rbind them into one data frame and save it.

```{r}
write.csv(DS, here("Output", "PostZooniverse", "Processed", "DS_workflow1.csv"))
write.csv(KA, here("Output", "PostZooniverse", "Processed", "KA_workflow1.1.csv"))
write.csv(CWMM, here("Output", "PostZooniverse", "Processed", "CWMM_speciesID.csv"))
```
And combine
```{r}
All<-rbind(DS,KA, CWMM)
All<-All %>% select(-X)
write.csv(All, here("Output", "PostZooniverse", "Processed", "all_data_to_June2022.csv"))
```

Ok, now let's gather some summary information about these files.
```{r}
All$created_at<-ymd_hms(All$created_at)
AllDataSummary<-All %>% group_by(Student) %>% summarize(
  NumObs=length(subject_ids),
  NumSubjects=length(unique(subject_ids)),
  NumClass=length(unique(classification_id)),
  min.class.date=min(created_at),
  max.class.date=max(created_at)
  
)
```

Ok! Now take these data and work on getting final classifications for the full data set.



