---
title: "Comparing datasets"
output: html_notebook
---

I'm curious to look at the range of dates for each of the 3 data sets I just made so I can see if I can decipher which projects are split between classification files.  

Begin by cleaning things up and pulling in data

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```
Pull in the data files

```{r}
DS<-read.csv(here("Output", "PostZooniverse", "Flattened", "north-country-wild-classifications.6June2022_flat_2022-06-07_wf1_donovan.csv"))

KA<-read.csv(here("Output", "PostZooniverse", "Flattened", "north-country-wild-classifications.6June2022_flat_2022-06-07_wf1.1_kate.csv"))

CWMM<-read.csv(here("Output", "PostZooniverse", "Flattened", "north-country-wild-classifications.6June2022_flat_2022-06-07_sp_id_cole_maggie.csv"))
```
Ok, now let's gather some summary information about these files.

Put a place to store the info

```{r}
myfiles<-c("WF1", "WF1.1", "WF_sp_id")
Meta<-data.frame(
  Workflow = character(3),
  Start.date = POSIXct(3),
  End.date = POSIXct(3),
  Class.ID.min = numeric(3),
  Class.ID.max = numeric(3)
)

Meta$Workflow<-myfiles
```

Now grab some of the information from each file:

```{r}
Meta$Start.date[1]<-min(DS$created_at)
Meta$End.date[1]<-max(DS$created_at)
Meta$Start.date[2]<-min(KA$created_at)
Meta$End.date[2]<-max(KA$created_at)
Meta$Start.date[3]<-min(CWMM$created_at)
Meta$End.date[3]<-max(CWMM$created_at)

Meta$Class.ID.min[1]<-min(DS$classification_id)
Meta$Class.ID.max[1]<-max(DS$classification_id)
Meta$Class.ID.min[2]<-min(KA$classification_id)
Meta$Class.ID.max[2]<-max(KA$classification_id)
Meta$Class.ID.min[3]<-min(CWMM$classification_id)
Meta$Class.ID.max[3]<-max(CWMM$classification_id)


```
Now get rid of the beta test classfications from CWMM and then save it.

```{r}
CWMM<-CWMM %>% filter(round != "beta test")
write.csv(CWMM, here("Output", "PostZooniverse", "Flattened", "no-co-wild-classifications.6June2022_flat_2022-06-07_sp_id_cole_maggie_no_beta_test.csv"))
```
How many unique subjects are there per file?

```{r}
Meta$num_subjects<-NA
Meta$num_subjects[1]<-length(unique(DS$subject_ids))
Meta$num_subjects[2]<-length(unique(KA$subject_ids))
Meta$num_subjects[3]<-length(unique(CWMM$subject_ids))
```
Create a CamSD column for each data set and use it to calculate the number of cameras in each dataset.

```{r}
DS$CamSD<- paste0(DS$CamNum,"-",DS$SD_card_num)
Meta$num_cams<-NA
Meta$num_cams[1]<-length(unique(DS$CamSD))

KA$CamSD<- paste0(KA$CamNum,"-",KA$SD_card_num)
Meta$num_cams[2]<-length(unique(KA$CamSD))

KA$CamSD<- paste0(KA$CamNum,"-",KA$SD_card_num)
Meta$num_cams[2]<-length(unique(KA$CamSD))

CWMM$CamSD<- paste0(CWMM$CamNum,"-",CWMM$SD_card_num)
Meta$num_cams[3]<-length(unique(CWMM$CamSD))
```
Now write meta, CWMM, DS and KA for further use.

```{r}
write.csv(DS, here("Code", "PostZooniverse", "2022-June", "Data", "DS.csv"))
write.csv(KA, here("Code", "PostZooniverse", "2022-June", "Data", "KA.csv"))
write.csv(CWMM, here("Code", "PostZooniverse", "2022-June", "Data", "CWMM.csv"))
write.csv(Meta, here("Code", "PostZooniverse", "2022-June", "Data", "Meta.csv"))
```


