---
title: "Exploring flattened data"
output: html_notebook
---

The end product of the flattening process is a new dataset that includes all of the species identifications and metadata associated with each event. We need to take those data and explore them to determine how well the classifications were made.

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```
Pull in data
```{r}
df<-read.csv(here("Output", "PostZooniverse","Processed","all_data_to_June2022.csv"))
```

What is the range in number of classifications per subject id?
```{r}
Class_summary<-df %>% group_by(Student, subject_ids) %>% summarise(
  NumClass=length(classification_id)
)

length(unique(df$subject_ids))
```
There are 15,344 subjects in the dataset

How many classifications?
```{r}
length(unique(df$classification_id))
```
There are a few classifications (83724-83387 = 337) that appear more than once. Why?

Let's look at them.

```{r}
df<-df %>% group_by(classification_id) %>% mutate(
  dupe = n()> 1)
ungroup(df)
check<-df %>% filter(dupe == TRUE)
```
So there are 337 cases in which there was > 1 animal observed in the subject. Cool. Looks good.


Now see if I can join the date-time of each object for each image from the camera exif data. I created a master DTO sheet to attempt with.

Read in the exif data
```{r}
exif<-read.csv(here("Data", "masterDTO.csv"))
exif$FileName2<-exif$FileName
```
Now try to join to df

```{r}
Main<-left_join(df, exif, by = c("Img1" = "FileName2"))
```
Now dump some columns to make it easier to look at
```{r}
Main<-Main %>% select(!c(FileSize, ImageSize, FocalLength, ShutterSpeed, Aperture, ISO, WhiteBalance, Flash))
```
And now relocate things
```{r}
Main<-Main %>% relocate(CamSD.y, .after = CamSD.x)
Main<-Main %>% relocate(DateTimeOriginal,.before = created_at)
Main<-Main %>% relocate(FileName, .before = user_name)
```

When I was working with this before, I was losing some of the info from maggie's data. Take a look:

```{r}
Main %>% filter(Student == "Maggie") %>% View()

```
Why is there one row in Maggie's data that looks like it is from a different dataset?  Check it out

```{r}
Main %>% filter(FileName == "14_22_DON_000031.JPG") %>% View()
```

For some reason it was classified again in Maggie's protocol - but we don't need it there, so I'm going to delete the one instance of it that corresponds to Maggie's data.
```{r}
find<-which(Main$FileName == "14_22_DON_000031.JPG" & Main$Student == "Maggie")
Main<-Main[-c(find),]
```
Now move workflow ID next to workflow version
```{r}
Main <-Main %>% relocate(workflow_id, .after = workflow_name)
```

Ok, now let's see if we can get all the forest names the same
```{r}
unique(Main$Forest)
```
Ok, let's do full forest names

```{r}
don<-which(Main$Forest == "DON")
Main$Forest[don]<-"Donnerville"
deg<-which(Main$Forest == "DEG")
Main$Forest[deg]<-"Degrasse"

bc<-which(Main$Forest == "BC")
Main$Forest[bc]<-"Beaver Creek"

sh<-which(Main$Forest == "SH")
Main$Forest[sh]<-"South Hammond"

wf<-which(Main$Forest == "WF")
Main$Forest[wf]<-"Whisky Flats"

whip<-which(Main$Forest == "WHIP" |Main$Forest == "Whip")
Main$Forest[whip]<-"Whippoorwill"

unique(Main$Forest)
```
Now fix round numbers
```{r}
unique(Main$round)
```
Make them all R1 or R2
```{r}
r1<-which(Main$round == "1")
Main$round[r1]<- "R1"

r2<-which(Main$round == "2")
Main$round[r2]<- "R2"

unique(Main$round)
```
Check more vars
```{r}
unique(Main$batch) #ok
unique(Main$CamNumber)
```
Find the cam nums that begin with "C"
```{r}
fix<-which(str_detect(Main$CamNumber, "C"))
View(Main[fix,])
#now pull the C off

Main$CamNumber[fix]<-str_replace(Main$CamNumber[fix], "C0", "0")
#now get rid of leading zeros
Main$CamNumber[fix]<-str_remove(Main$CamNumber[fix], "^0+")
Main$CamNumber<-as.numeric(Main$CamNumber)
Main$CamSD.z<-paste0(Main$CamNumber,"-", Main$SDNumber)
Main<-Main %>% relocate(CamSD.z, .after = CamSD.y)
```
Now forest type
```{r}
unique(Main$For_type)
fixpines<-which(Main$For_type == "Pine ")
Main$For_type[fixpines]<-"Pine"
fixdecid<-which(Main$For_type == "Deciduous ")
Main$For_type[fixdecid]<-"Deciduous"
unique(Main$For_type)
```
Fix phase - combine phase 1 batch 1 and phase 1 batch 2 as "experiment"

```{r}
exp<-which(Main$Phase == "phase 1 batch 1"| Main$Phase == "phase 1 batch 2")

Main$Phase[exp]<-"experiment"
unique(Main$Phase)
```
check Treatment
```{r}
unique(Main$Treatment)
```
Need to combine Snap and Snapshot
```{r}
snap<-which(Main$Treatment == "Snap")
Main$Treatment[snap]<-"Snapshot"
unique(Main$Treatment)
```
Now check Round
```{r}
unique(Main$Round)
```
That's fine.
Now move Cam_SD next to CamSD.z, Cam and SD after CamNumber, SDNumber
```{r}
Main<-Main %>% relocate(Cam_SD, .after = CamSD.z)
Main<-Main %>% relocate(Cam, .after = SDNumber) %>% relocate(SD, .after = Cam)
Main<-Main %>% relocate(Round, .after = round) %>% relocate(Year, .after = DTO) %>% relocate(Model, .after =CamMod) %>% relocate(dupe, .after = retire.date)
Main<-Main %>% select(-X)
```












