---
title: "Consensus picks"
output: html_notebook
---

Once we have flattened all of our data, we need to examine it to determine the consensus species ID for each event.

We need to know the following:

1. For how many subject ids were all of the choices the same (i.e. all classifiers identified the same species)?
  * What was the chosen species for each of these subject ids?


2. For which subject_ids were there discrepancies in what the species_id was?
  * How many of these subject_ids were there?

3. For which subject_ids were there more than one species present in the set of photos?
  * We'll see below that these are the cases in which num_votes > num_class
  * Are there discrepancies in what species are indicated as present? How will we reconcile them?

## Set up the workspace and read in the data
```{r}
rm(list = ls())
library(tidyverse)
library(here)
```
Pull in the master classification file

```{r}
df<-read.csv(here("Output", "PostZooniverse", "Processed", "master_class_file_to_June2022.csv"))
df<-df %>% select(-X)
```


Summarize the data to see the number of distinct subjects (events) and the number of distinct classifications in the data set. 
```{r}
df %>% summarise(n_distinct(new_sub_id), n_distinct(classification_id)) 
```
Create a new column that has a count of how many unique classifications were made.

```{r}
df<-df %>% 
     group_by(new_sub_id) %>% # count up the number of distinct classification IDs
     mutate( 
            num_class = n_distinct(classification_id),
            num_diff_species = n_distinct(choice)) %>% #because there will be >1 row per classification_id if >1 spp
     arrange(new_sub_id, classification_id) 
df %>% View
unique(df$num_diff_species)
unique(df$num_class)
```
Some were classified over and over again, which is fine.


Now arrange in descending order of the number of species detected.

```{r} 
df %>% arrange(desc(num_diff_species), new_sub_id) %>% View()
```
Let's look at a histogram of that

```{r}
ggplot(df, aes(num_diff_species))+
  geom_histogram()
```
Good - mostly 1 or 2 species detected.

Now that we know the number of species identified per subject id, we need a way to  determine the consensus species for each subject id. First we'll write some code to check the species counts per subject id and classification id.

## Subject-level metrics
Generate some information about each subject_id
```{r}
df <- df %>% ungroup %>%
     group_by(new_sub_id) %>%
     mutate(num_votes = n(), # if a  user ids >1 spp, there will be more votes than classifications
            ) 
#range in number of votes
unique(df$num_votes)
```


Find the rows for which the number of votes is 2x number of classifications
```{r}
double_votes<-df %>% filter(num_votes == 2*num_class)
```
All the double_votes are cases where more than one species was detected in the subject set - which is good.

Now look at species-level metrics

```{r}
species_counts <- df %>% ungroup %>%
     group_by(new_sub_id, classification_id) %>%
     summarise(
       total_spp_by_user = mean(num_diff_species)) %>% #Need to select only one row per classification_id, then summarise across those. 
     summarise(., agg_num_species = round(median(total_spp_by_user), 0))#aggregate species count, which is median rounded up
     glimpse(species_counts)
     
#species counts gives a list by subject_id of the number of species identified by the different classifiers.
```
Now get the cleaned classifications
```{r}
cleaned_classifications <- left_join(df, species_counts) %>% ungroup
glimpse(cleaned_classifications)
```

## Species level metrics
```{r}
grouped_classifications <- cleaned_classifications %>% 
      group_by(new_sub_id, num_class, num_votes, agg_num_species, num_diff_species, choice) # fields at subject level or higher
```

#Tally the votes for each species ID'd within a subject
This yields species_votes, which tells us, for each subject_id, the number of classifications and the proportion of the vote received by each species for each subject_id. We can use this to run a list of subjects for which we need to go back and confirm the species.
```{r}
species_votes <- grouped_classifications %>% 
     # for every species within a subject, aggregate votes.
     summarise(., votes = n_distinct(classification_id),
               ) %>% #count up the number of votes per species choice
     mutate(propvote = votes/sum(votes), #calculate proportion of votes for this species
            propclass = votes/num_class) 

#propvote will not equal propclass when the total number of votes is different than the total number of classifications. The number of votes exceeds the number of classifications when there is more than one species in a subject set (e.g. dog + human.)
```
When num_votes > num_class you have a situation where there was more than one species detected in an event (typically Gus and one or more camera trappers.). In these cases, propvote will not equal propclass, and the sum of propclass for that subject id will be greater than 1.0 (though the sum of propvote will be equal to 1.)

```{r}
MoreThan1<-species_votes %>% filter(num_votes > num_class)
```
They each show the same subject_ids so things look good.

Let's filter species_votes to extract the subjects for which there is more than 1 species but the number of votes = num_class. These will be the cases in which there is one species in the event but there is a discrepancy about what the species is.

```{r}
subjects_to_check<-species_votes %>% filter(num_votes == num_class & num_diff_species > 1)

```
Now let's look at the distribution of classifications among subjects. 

```{r}
summary_class<-species_votes %>% ungroup %>% group_by(new_sub_id) %>% summarise(
  num_class = max(num_class)
)

ggplot(summary_class, aes(num_class))+
  geom_histogram()

table(summary_class$num_class)
```
There are 30 classifications that were only looked at once or twice, and an additional 511 that were only looked at 3 times. I should go through all of those by hand. And perhaps also the 553 that were looked at only 4 times. Let's set them aside in a list

```{r}
too_few<-filter(summary_class, num_class <= 4)
not_classed<-species_votes %>% filter(new_sub_id %in% too_few$new_sub_id)
```

Now we want to filter out the cases for which propvote = propclass and propclass is 0.8 or greater or 0.2 or smaller. But we don't want to include the cases in which there weren't enough classifications (which would have an inflated propvote/propclass, e.g. if there is only 1 classfication, propvote = 1.0)

```{r}
#step 1 - Remove the not_classed from subjects_to_check
do_not_include<-which(subjects_to_check$new_sub_id %in% not_classed$new_sub_id)
classed_to_check<-subjects_to_check[-do_not_include,]
#step 2 - filter out those cases with high or low propvote
species_votes_to_check<-classed_to_check %>% filter((propvote == propclass) & (between(propclass, 0.21, 0.79)))

```

now combine not_classed with species_votes_to_check and then remove duplicates.

```{r}
need_checking<-rbind(not_classed, species_votes_to_check)
#remove duplicates
need_checking <- distinct(need_checking)
```

Now get the list of the subject ids in need_checking and use it to filter cleaned_classifications to get the list of subjects, images, events etc. to check.

```{r}
subjects_to_filter<-unique(need_checking$new_sub_id)
```


And now filter cleaned_classifications by those data

```{r}
filesToCheck<-filter(cleaned_classifications, new_sub_id %in% subjects_to_filter)
```
And now summarize
```{r}
filesToCheck<-filesToCheck %>% group_by(new_sub_id) %>% summarize(
  workflow_name = workflow_name,
  workflow_version = workflow_version,
  Img1 = Img1,
  Img2 = Img2,
  Img3 = Img3,
  DateTimeOriginal = DateTimeOriginal,
  round = round,
  Round = Round,
  Forest = Forest,
  CamNumber = CamNumber,
  SDNumber = SDNumber,
  Cam = Cam,
  SD = SD,
  Year = Year,
  Student = Student
)
```
Now loop through filesToCheck to grab just the first instance of each subject id and associated variables.  

Make data frame to put results in 
```{r}
Subjects<-unique(filesToCheck$new_sub_id)

Checklist<-data.frame(new_sub_id = character(length(Subjects)),
                      workflow_name = character(length(Subjects)),
                      workflow_version = numeric(length(Subjects)),
                      Img1 = character(length(Subjects)), 
                      Img2 = character(length(Subjects)), 
                      Img3 = character(length(Subjects)),
                      DateTimeOriginal = character(length(Subjects)),
                      round = character(length(Subjects)),
                      Round = character(length(Subjects)),
                      Forest = character(length(Subjects)),
                      CamNumber = integer(length(Subjects)),
                      SDNumber = integer(length(Subjects)),
                      Cam = integer(length(Subjects)),
                      SD = integer(length(Subjects)),
                      Year = integer(length(Subjects)),
                      Student = character(length(Subjects)))
```
Now loop
```{r}
for(i in 1:length(Subjects)){
  mySample<-filter(filesToCheck, new_sub_id == Subjects[i])
  Checklist$new_sub_id[i]<-Subjects[i]
  Checklist$workflow_name[i]<-unique(mySample$workflow_name)
  Checklist$workflow_version[i]<-unique(mySample$workflow_version)
  Checklist$Img1[i]<-unique(mySample$Img1)
  Checklist$Img2[i]<-unique(mySample$Img2)
  Checklist$Img3[i]<-unique(mySample$Img3)
  Checklist$DateTimeOriginal[i]<-unique(mySample$DateTimeOriginal)
  Checklist$round[i]<-unique(mySample$round)
  Checklist$Round[i]<-unique(mySample$Round)
  Checklist$Forest[i]<-unique(mySample$Forest)
  Checklist$CamNumber[i]<-unique(mySample$CamNumber)
  Checklist$SDNumber[i]<-unique(mySample$SDNumber)
  Checklist$Cam[i]<-unique(mySample$Cam)
  Checklist$SD[i]<-unique(mySample$SD)
  Checklist$Year[i]<-unique(mySample$Year)
  Checklist$Student[i]<-unique(mySample$Student)
}
```
Add column for assigning the choice
```{r}
Checklist$choice<-NA
Checklist<-Checklist %>% relocate(choice, .after = Img3)
```

Now save

```{r}
write.csv(Checklist, here("Output", "PostZooniverse", "master_files_to_checkJune2022.csv"))
```

Now we need to assign the correct species for the files that we are not checking:
Extract from cleaned classifications all of the subject ids that are NOT part of the list that needs to be checked:

```{r}
Subjects_to_keep<-species_votes %>% filter(propvote >= 0.8 & num_class > 4)
```
Now see if there are any of the subject ids from Checklist left in Subjects_to_keep

```{r}
check<-which(Subjects_to_keep$new_sub_id %in% Checklist$new_sub_id)
```
No - so that's good!


```{r}
NeedAssignment<- cleaned_classifications %>% filter(new_sub_id %in% Subjects_to_keep$new_sub_id)


NeedAssignSummary<-NeedAssignment %>% group_by(new_sub_id) %>% summarise(
  num_class = n_distinct(classification_id),
  num_species = n_distinct(choice)
)
```
A subset have more than 1 species - these are likely the instances in which there were > 1 species in the event. Let's take a look.


```{r}
check<-NeedAssignSummary %>% filter(num_species > 1)
MoreThan1<-NeedAssignment %>% filter(new_sub_id %in% check$new_sub_id)
```
Many have more than 1, but they all have a majority of one species, so we want to choose that species. We could order the data by subject_id and then prop_vote(if we connect it) and then choose the one for which propvote > 0.8.

Let's see.

```{r}
NeedAssignment<-left_join(NeedAssignment, Subjects_to_keep)
NeedAssignment<-NeedAssignment %>% arrange(new_sub_id, propvote)
```


Now let's clean up the environment - overwhelming
```{r}
rm(double_votes, grouped_classifications, MoreThan1, mySample, need_checking, not_classed, species_counts, species_votes, species_votes_to_check, subjects_to_check, summary_class, too_few, Subjects, subjects_to_filter)

```

Now create data frame to accept final assignments

```{r}
how_long<-n_distinct(NeedAssignment$new_sub_id)
Assigned<-data.frame(new_sub_id = character(how_long),
                     workflow_name = character(how_long),
                     workflow_version = numeric(how_long),
                      Img1 = character(how_long), 
                      Img2 = character(how_long), 
                      Img3 = character(how_long),
                      choice = character(how_long),
                     DateTimeOriginal = character(how_long),
                      round = character(how_long),
                      Round = character(how_long),
                     Forest = character(how_long),
                      CamNumber = integer(how_long),
                      SDNumber = integer(how_long),
                      Cam = integer(how_long),
                      SD = integer(how_long),
                      Year = integer(how_long),
                      Student = character(how_long))

```
Now loop
```{r}
Subjects<-unique(NeedAssignment$new_sub_id)
for(i in 1:how_long){
  mySample<-filter(NeedAssignment, new_sub_id == Subjects[i])
  Assigned$new_sub_id[i]<-Subjects[i]
  Assigned$workflow_name[i]<-unique(mySample$workflow_name)
  Assigned$workflow_version[i]<-unique(mySample$workflow_version)
  Assigned$Img1[i]<-unique(mySample$Img1)
  Assigned$Img2[i]<-unique(mySample$Img2)
  Assigned$Img3[i]<-unique(mySample$Img3)
  Assigned$choice[i]<-unique(mySample$choice[1])
  Assigned$DateTimeOriginal[i]<-unique(mySample$DateTimeOriginal)
  Assigned$round[i]<-unique(mySample$round)
  Assigned$Round[i]<-unique(mySample$Round)
  Assigned$Forest[i]<- unique(mySample$Forest)
  Assigned$CamNumber[i]<-unique(mySample$CamNumber)
  Assigned$SDNumber[i]<-unique(mySample$SDNumber)
  Assigned$Cam[i]<-unique(mySample$Cam)
  Assigned$SD[i]<-unique(mySample$SD)
  Assigned$Year[i]<-unique(mySample$Year)
  Assigned$Student[i]<-unique(mySample$Student)
  
  
  
 
}
```
Now save

```{r}
write.csv(Assigned, here("Output", "PostZooniverse","master_assigned_AssignedSpecies_June2022.csv"))
```

I notice that there are 13315 subjects that have species assigned and 1983 subjects that need to be checked. That adds to 15298. There are a total of 15390 subject ids, so I'm not sure why we are missing 92 from all of this. Let's see which once are missing.

```{r}
subs1<-Assigned$new_sub_id
subs2<-Checklist$new_sub_id
all_subs<-c(subs1,subs2)
my92<-setdiff(unique(df$new_sub_id), all_subs)
see92<-cleaned_classifications %>% filter(new_sub_id %in% my92)
see92<-see92 %>% arrange(new_sub_id, classification_id, array.index)
```
Now look at the ones for which num_votes = num_class.

```{r}
see92 %>% filter(num_votes == num_class) %>% View()
```
All are cases where there are as many species choices as there are votes. We need to add this set to Checklist.

Create checklist 2
```{r}
check2<-see92 %>% filter(num_votes == num_class)
#summarise

check2<-check2 %>% group_by(new_sub_id) %>% summarize(
  workflow_name = workflow_name,
  workflow_version = workflow_version,
  Img1 = Img1,
  Img2 = Img2,
  Img3 = Img3,
  DateTimeOriginal = DateTimeOriginal,
  round = round,
  Round = Round,
  Forest = Forest,
  CamNumber = CamNumber,
  SDNumber = SDNumber,
  Cam = Cam,
  SD = SD,
  Year = Year,
  Student = Student
)
```
Now loop through check2 to grab just the first instance of each subject id and associated variables.  

Make data frame to put results in 
```{r}
Subjects<-unique(check2$new_sub_id)

Checklist2<-data.frame(new_sub_id = character(length(Subjects)),
                      workflow_name = character(length(Subjects)),
                      workflow_version = numeric(length(Subjects)),
                      Img1 = character(length(Subjects)), 
                      Img2 = character(length(Subjects)), 
                      Img3 = character(length(Subjects)),
                      DateTimeOriginal = character(length(Subjects)),
                      round = character(length(Subjects)),
                      Round = character(length(Subjects)),
                      Forest = character(length(Subjects)),
                      CamNumber = integer(length(Subjects)),
                      SDNumber = integer(length(Subjects)),
                      Cam = integer(length(Subjects)),
                      SD = integer(length(Subjects)),
                      Year = integer(length(Subjects)),
                      Student = character(length(Subjects)))
```
Now loop
```{r}
for(i in 1:length(Subjects)){
  mySample<-filter(check2, new_sub_id == Subjects[i])
  Checklist2$new_sub_id[i]<-Subjects[i]
  Checklist2$workflow_name[i]<-unique(mySample$workflow_name)
  Checklist2$workflow_version[i]<-unique(mySample$workflow_version)
  Checklist2$Img1[i]<-unique(mySample$Img1)
  Checklist2$Img2[i]<-unique(mySample$Img2)
  Checklist2$Img3[i]<-unique(mySample$Img3)
  Checklist2$DateTimeOriginal[i]<-unique(mySample$DateTimeOriginal)
  Checklist2$round[i]<-unique(mySample$round)
  Checklist2$Round[i]<-unique(mySample$Round)
  Checklist2$Forest[i]<-unique(mySample$Forest)
  Checklist2$CamNumber[i]<-unique(mySample$CamNumber)
  Checklist2$SDNumber[i]<-unique(mySample$SDNumber)
  Checklist2$Cam[i]<-unique(mySample$Cam)
  Checklist2$SD[i]<-unique(mySample$SD)
  Checklist2$Year[i]<-unique(mySample$Year)
  Checklist2$Student[i]<-unique(mySample$Student)
}
```
Add column for assigning the choice
```{r}
Checklist2$choice<-NA
Checklist2<-Checklist2 %>% relocate(choice, .after = Img3)
```
Add them to Checklist
```{r}
FinalCheck<-rbind(Checklist, Checklist2)
write.csv(FinalCheck, here("Output", "PostZooniverse", "master_files_to_checkJune2022.csv"))
```



Now deal with those that have multiple species per subject set.

```{r}
MoreThan<-see92 %>% filter(num_votes != num_class) 
```
I also want to just double check these, so let me add them as well.

Create checklist 3
```{r}
check3<-MoreThan
#summarise

check3<-check3 %>% group_by(new_sub_id) %>% summarize(
  workflow_name = workflow_name,
  workflow_version = workflow_version,
  Img1 = Img1,
  Img2 = Img2,
  Img3 = Img3,
  DateTimeOriginal = DateTimeOriginal,
  round = round,
  Round = Round,
  Forest = Forest,
  CamNumber = CamNumber,
  SDNumber = SDNumber,
  Cam = Cam,
  SD = SD,
  Year = Year,
  Student = Student
)
```
Now loop through check3 to grab just the first instance of each subject id and associated variables.  

Make data frame to put results in 
```{r}
Subjects<-unique(check3$new_sub_id)

Checklist3<-data.frame(new_sub_id = character(length(Subjects)),
                      workflow_name = character(length(Subjects)),
                      workflow_version = numeric(length(Subjects)),
                      Img1 = character(length(Subjects)), 
                      Img2 = character(length(Subjects)), 
                      Img3 = character(length(Subjects)),
                      DateTimeOriginal = character(length(Subjects)),
                      round = character(length(Subjects)),
                      Round = character(length(Subjects)),
                      Forest = character(length(Subjects)),
                      CamNumber = integer(length(Subjects)),
                      SDNumber = integer(length(Subjects)),
                      Cam = integer(length(Subjects)),
                      SD = integer(length(Subjects)),
                      Year = integer(length(Subjects)),
                      Student = character(length(Subjects)))
```
Now loop
```{r}
for(i in 1:length(Subjects)){
  mySample<-filter(check3, new_sub_id == Subjects[i])
  Checklist3$new_sub_id[i]<-Subjects[i]
  Checklist3$workflow_name[i]<-unique(mySample$workflow_name)
  Checklist3$workflow_version[i]<-unique(mySample$workflow_version)
  Checklist3$Img1[i]<-unique(mySample$Img1)
  Checklist3$Img2[i]<-unique(mySample$Img2)
  Checklist3$Img3[i]<-unique(mySample$Img3)
  Checklist3$DateTimeOriginal[i]<-unique(mySample$DateTimeOriginal)
  Checklist3$round[i]<-unique(mySample$round)
  Checklist3$Round[i]<-unique(mySample$Round)
  Checklist3$Forest[i]<-unique(mySample$Forest)
  Checklist3$CamNumber[i]<-unique(mySample$CamNumber)
  Checklist3$SDNumber[i]<-unique(mySample$SDNumber)
  Checklist3$Cam[i]<-unique(mySample$Cam)
  Checklist3$SD[i]<-unique(mySample$SD)
  Checklist3$Year[i]<-unique(mySample$Year)
  Checklist3$Student[i]<-unique(mySample$Student)
}
```
Add column for assigning the choice
```{r}
Checklist3$choice<-NA
Checklist3<-Checklist2 %>% relocate(choice, .after = Img3)
```

And add to final check
```{r}
FinalCheck<-rbind(FinalCheck, Checklist3)
write.csv(FinalCheck, here("Output", "PostZooniverse", "master_files_to_checkJune2022.csv"))
```

Now I have 2005 observations in FinalCheck and 13315 in Assigned = 15320

So I'm still missing 70 observations...

```{r}
still_not_there<-setdiff(my92, FinalCheck$new_sub_id)
```

