---
title: "Merging-flat-files"
author: "Erika Barthelmess"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I'm working on generating a master flat file that includes all classifications through 13 June 2023.  I have taken the three flat files with data that I generated yesterday:
- wkfl_9599_123.17_flat_2023-06-12.csv
- wkfl_12161_6.5_flat_2023-06-12.csv
- wkfl_16695_29.1_flat_2023-06-12.csv

and have adjusted them in excel so that they all have the same columns in the same order. I saved those adjusted files as:
- wkfl_9599_123.17_flat_col.csv
- wkfl_12161_6.5_flat_col.csv
- wkfl_16695_29.1_flat_col.csv

Now I'm going to put the data from these three files together into one master file for further processing.

Set things up:
```{r}
rm(list =ls())
library(tidyverse)
library(here)
```

Read in the files
```{r}
DF1 <- read.csv(here("Output", "PostZooniverse", "Flattened", "wkfl_9599_123.17_flat_cols.csv"))
DF2 <-read.csv(here("Output", "PostZooniverse", "Flattened", "wkfl_12161_6.5_flat_cols.csv"))
DF3<- read.csv(here("Output", "PostZooniverse", "Flattened", "wkfl_16695_29.1_flat_cols.csv"))
```
Now check to see that colnames match in each DF
```{r}
colnames(DF1)==colnames(DF2)
colnames(DF1)==colnames(DF3)
```
Good! Now I can make one master DF





