---
title: "Merging-flat-files"
author: "Erika Barthelmess"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I'm working on generating a master flat file that includes all classifications through 13 June 2023.  I have taken the three flat files with data that I generated yesterday:
- wkfl_9599_123.17_flat_2023-06-12.csv
- wkfl_12161_6.5_flat_2023-06-12.csv
- wkfl_16695_29.1_flat_2023-06-12.csv

and have adjusted them in excel so that they all have the same columns in the same order. I saved those adjusted files as:
- wkfl_9599_123.17_flat_col.csv
- wkfl_12161_6.5_flat_col.csv
- wkfl_16695_29.1_flat_col.csv

Now I'm going to put the data from these three files together into one master file for further processing.

Set things up:
```{r}
rm(list =ls())
library(tidyverse)
library(here)
```

Read in the files
```{r}
DF1 <- read.csv(here("Output", "PostZooniverse", "Flattened", "wkfl_9599_123.17_flat_cols.csv"))
DF2 <-read.csv(here("Output", "PostZooniverse", "Flattened", "wkfl_12161_6.5_flat_cols.csv"))
DF3<- read.csv(here("Output", "PostZooniverse", "Flattened", "wkfl_16695_29.1_flat_cols.csv"))
```
Now check to see that colnames match in each DF
```{r}
colnames(DF1)==colnames(DF2)
colnames(DF1)==colnames(DF3)
```
Good! Now I can make one master DF

```{r}
Master <- rbind(DF1, DF2, DF3)
```
Now take a look at the data in Master
```{r}
summary(Master)
```

Some of the rows for the FILE column are empty. Let's take a look

```{r}
check<-subset(Master, is.na(FILE)| trimws(FILE)=="")
unique(check$workflow_name)
unique(check$workflow_version)
```
So all of the blanks are from the same file. Need to replace that information into the Master.

```{r}
fix<-which(is.na(Master$FILE)|trimws(Master$FILE)=="")
Master$FILE[fix]<-"wkfl_12161_6.5_flat_2023-06-12"
```

All better.  Now save the file.

```{r}
write.csv(Master, here("Output", "PostZooniverse", "Flattened", "flat_through_2023-06-13.csv"))
```

