---
title: "Determine Zooniverse Leaderboard"
output: html_notebook
---

This file steps the user through opening a "raw" zooniverse classification file and extracting data from your choice of workflows.

Set things up
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(lubridate)
library(rstudioapi)

myFunctions<-list.files(here("Code", "Functions"), pattern = "R") 
#points to files that need to be sourced

sapply(here("Code", "Functions", myFunctions), source) #sources all of the functions in the Functions folder.
```
###Step 1. Choose the classification file to work with.
We have multiple classification files. The `choose_my_classification_file` function asks the user for input so that they can select the classification file to work with. The classification file should be on the t://drive in Bart_s04/Research_group/NoCoWild/Data/classification_files.

```{r}
choose_class_file<-function(){
  require(rstudioapi) #turns on Rstudio api if not alreay loaded
  #choose classification file; requires active t:drive connection
  rstudioapi::showDialog(
    title = "Let's give R some file information",
    message = "Select the file containing the classification data you wish to process")

  #path to exif file folder on t:drive
  path_to_class_file<-"/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/classification_files/"

  class_file_path<-rstudioapi::selectFile(
    path = "/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/classification_files/",
    filter = "CSV files (*.csv)",
    label= "Select desired classification file"
  )

  jdata<-read.csv(class_file_path)
  filename<-basename(class_file_path)
  output<-list(jdata, filename)
}
```

Now run the function
```{r}
data<-choose_class_file()
df<-data[[1]]
```
Now isolate the the workflow(s) you are interested in, saving each as a list-element that is a df

```{r}
get_workflows<-function(df){
  require(dplyr)
  #identify workflows by making workflow_id into a factor
  workflow_nums<-unique(df$workflow_id)
  workflow_names<-unique(df$workflow_name)
  workflows<-paste0(workflow_nums, " - ", workflow_names)

  #now ask user which workflow or workflows to use
  my_workflows <- select.list(workflows , multiple=TRUE,
                      title='Which workflow(s) do you want to extract?')
    #now filter the selections
    #first generate workflow ids from the selection
  my_workflow_positions<-which(workflows %in% my_workflows)
  my_workflow_nums<-workflow_nums[my_workflow_positions]
  
  n_workflows<-length(my_workflow_nums)
  
  workflow_versions<-
  #extract these workflows
  extracted_workflows<-list()
  for(i in 1:n_workflows){
    temp<-df %>% filter(workflow_id == my_workflow_nums[i])
    extracted_workflows[[i]]<-temp
  }
  
  #now ask user which workflow version to use for each workflow
  
 
  
  

  DF2<-filter(df, cYear %in% my_years)
  my_months <-select.list(unique(DF2$cMonth), multiple = TRUE,
                          title = "In what month(s) where the classifications you are after made?")

  DF2<-filter(DF2, cMonth %in% my_months)
  ############
  
  my_selection<- menu(workflows,
                      title= "Which workflow do you want?")

  my_selection_filtered<-workflow_nums[my_selection]
  #now find the desired version of the workflow
  DF<-filter(jdata, workflow_id == my_selection_filtered)
  DF$workflow_version<-as.factor(
    DF$workflow_version)

  versions<-levels(DF$workflow_version)

  my_wkflow_version<- menu(
    versions,
    title = "Which workflow version do you want?")
  #now filter df to that workflow version

  DF2<-filter(
    DF,
    workflow_version == versions[my_wkflow_version])
out<-list(DF2, filename)
  return(out)
}
Pull in data
Use choose file dialog
```{r}
#choose classification file; requires active t:drive connection
rstudioapi::showDialog(
  title = "Let's give R some file information",
  message = "Select the file containing the flattened classification data you wish to process")

#path to exif file folder on t:drive
path_to_flat_file<-"/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/classification_files/"

flat_file_path<-rstudioapi::selectFile(
  path = "/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/classification_files/",
  filter = "CSV files (*.csv)",
  label= "Select desired flat file"
)

Classified<-read.csv(flat_file_path)
filename<-basename(flat_file_path)
}
```


  #now isolate workflow; code follows the choose_my_workflow function
  require(dplyr)
  #identify workflows by making workflow_id into a factor
  workflow_nums<-unique(jdata$workflow_id)
  workflow_names<-unique(jdata$workflow_name)
  workflows<-paste0(workflow_nums, " - ", workflow_names)

  #now ask user which workflow to use

  my_selection<- menu(workflows,
                      title= "Which workflow do you want?")

  my_selection_filtered<-workflow_nums[my_selection]
  #now find the desired version of the workflow
  DF<-filter(jdata, workflow_id == my_selection_filtered)
  DF$workflow_version<-as.factor(
    DF$workflow_version)

  versions<-levels(DF$workflow_version)

  my_wkflow_version<- menu(
    versions,
    title = "Which workflow version do you want?")
  #now filter df to that workflow version

  DF2<-filter(
    DF,
    workflow_version == versions[my_wkflow_version])
out<-list(DF2, filename)
  return(out)
}
Pull in data
Use choose file dialog
```{r}
#choose classification file; requires active t:drive connection
rstudioapi::showDialog(
  title = "Let's give R some file information",
  message = "Select the file containing the flattened classification data you wish to process")

#path to exif file folder on t:drive
path_to_flat_file<-"/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/classification_files/"

flat_file_path<-rstudioapi::selectFile(
  path = "/Volumes/classes/Bart_s04/Research_group/NoCoWild/Data/classification_files/",
  filter = "CSV files (*.csv)",
  label= "Select desired flat file"
)

Classified<-read.csv(flat_file_path)
filename<-basename(flat_file_path)
```
