---
title: "Consensus picks"
output: html_notebook
---

Once we have flattened all of our data, we need to examine it to determine the consensus species ID for each event.

We need to know the following:

1. For how many subject ids were all of the choices the same (i.e. all classifiers identified the same species)?
  * What was the chosen species for each of these subject ids?

2. For which subject_ids were there discrepancies in what the species_id was?
  * How many of these subject_ids were there?

3. For which subject_ids were there more than one species present in the set of photos?
  * We'll see below that these are the cases in which num_votes > num_class
  * Are there discrepancies in what species are indicated as present? How will we reconcile them?

## Set up the workspace and read in the data
```{r, results ='hide'}
rm(list = ls())
library(tidyverse)
library(here)
```
Set up the path to the flattened classification file

```{r}
rstudioapi::showDialog(
  title = "Let's give R some file information",
  message = "Select the file containing the flattened classfication file you wish to use")

#path to flattened file 
path_to_flat_file<-here("Output", "PostZooniverse", "Flattened")

flat_file_path<-rstudioapi::selectFile(
  path = path_to_flat_file,
  filter = "CSV files (*.csv)",
  label= "Select desired flattened classification file"
)

Flat <- read.csv(flat_file_path) #obviously, put correct file name here

```
Manage and decompose file name for later use

```{r}
file_name <-basename(flat_file_path)
file_name_parts<-unlist(strsplit(file_name, "_"))
```


Check to see if data are from multiple rounds. If beta test round is present, get rid of it
```{r}
unique(Flat$round)
```


```{r}
#get rid of data from beta test
#Flat<-Flat %>% filter(round !="beta test")
```

Parse the flat file into several data frames for simplicity; they will be joined back together later.

1. Get rid of junk columns added by writing csv files (e.g. "X", "X.1")
```{r}
#use select and ! (not) to exclude columns that start with X
Flat <- Flat %>% select(!starts_with("X"))
```
Now for ease of viewing, get rid of all columns that are empty. First determine which they are
```{r}

empty_cols<-sapply(Flat, function(x)all(is.na(x))) #identify the columns that only contain NA

#use dplyr to get rid of them. Note that if the count of NAs is equal to the number of rows, the column must be all NAs

Flat <- Flat %>%  select(
  where(
    ~sum(!is.na(.x)) > 0
  )
)
#now add back Event column
Flat$Event <- NA

```


2.Summarize the data to see the number of distinct subjects (events) and the number of distinct classifications in the data set. 
```{r}
Flat %>% summarise(n_distinct(subject_ids), n_distinct(classification_id)) 
```
Create a new column that has a count of how many unique classifications were made.

```{r}
Flat<-Flat %>% 
     group_by(subject_ids) %>% # count up the number of distinct classification IDs
     mutate( 
            num_class = n_distinct(classification_id),
            num_diff_species = n_distinct(choice)) %>% #because there will be >1 row per classification_id if >1 spp
     arrange(subject_ids, classification_id) 
Flat %>% View
```

Now arrange in descending order of the number of species identified.

```{r} 
Flat %>% arrange(desc(num_diff_species)) %>% View()
```

Now that we know the number of species identified per subject id, we need a way to  determine the consensus species for each subject id. First we'll write some code to check the species counts per subject id and classification id.

## Subject-level metrics
Generate some information about each subject_id
```{r}
subject_metrics <- Flat %>% ungroup %>%
     group_by(subject_ids) %>%
     mutate(num_votes = n(), # if a  user ids >1 spp in a classification (e.g. dog and human), there will be more votes than classifications
            ) 
```
Now look at species-level metrics

```{r}
species_counts <- Flat %>% ungroup %>%
     group_by(subject_ids, classification_id) %>%
     summarise(total_spp_by_user = mean(num_diff_species)) %>% #Need to select only one row per classification_id, then summarise across those. 
     summarise(., agg_num_species = round(median(total_spp_by_user), 0))#aggregate species count, which is median rounded up
     glimpse(species_counts)
     
#species counts gives a list by subject_id of the number of species identified by the different classifiers.
```
Now get the cleaned classifications
```{r}
cleaned_classifications <- left_join(subject_metrics, species_counts) %>% ungroup
glimpse(cleaned_classifications)
```

## Species level metrics
```{r}
grouped_classifications <- cleaned_classifications %>% 
      group_by(subject_ids, num_class, num_votes, agg_num_species, num_diff_species, choice) # fields at subject level or higher
```

#Tally the votes for each species ID'd within a subject
This yields species_votes, which tells us, for each subject_id, the number of classifications and the proportion of the vote received by each species for each subject_id. We can use this to run a list of subjects for which we need to go back and confirm the species.
```{r}
species_votes <- grouped_classifications %>% 
     # for every species within a subject, aggregate votes.
     summarise(., votes = n_distinct(classification_id),
               Event = max(Event)) %>% #count up the number of votes per species choice
     mutate(propvote = votes/sum(votes), #calculate proportion of votes for this species
            propclass = votes/num_class) 
```
Now obtain the correct classifications. These are the cases for which the number of classifications = number of votes AND for which the num_diff_species = 1. In other words, for all votes cast, there was agreement on what the subject contained.

```{r}
correct_classifications <- cleaned_classifications %>% filter(num_class == num_votes & num_diff_species == 1)
#now get list of correct subject ids

correct_subjects <- unique(correct_classifications$subject_ids)
```

When num_votes > num_class you have a situation where there was more than one species detected in an event (typically Gus and one or more camera trappers.). In these cases, propvote will not equal propclass, and the sum of propclass for that subject id will be greater than 1.0 (though the sum of propvote will be equal to 1.)

```{r}
MoreThan1<-species_votes %>% filter(num_votes > num_class)
```
They each show the same subject_ids so things look good.

Let's filter species_votes to extract the subjects for which there is more than 1 species but the number of votes = num_class. These will be the cases in which there is one species in the event but there is a discrepancy about what the species is.

```{r}
subjects_to_check<-species_votes %>% filter(num_votes == num_class & num_diff_species > 1)

```
Now we want to filter out the cases for which propvote = propclass and propclass is 0.8 or greater or 0.2 or smaller.

```{r}
species_votes_to_check<- subjects_to_check %>% filter((propvote == propclass) & (between(propclass, 0.21, 0.79)))
```
Now get the list of the subject ids in species_votes_to_check and use it to filter cleaned_classifications to get the list of subjects, images, events etc. to check.

```{r}
subjects_to_filter<-unique(species_votes_to_check$subject_ids)
```


And now filter cleaned_classifications by those data and by the correct data

```{r}
correct_classifications <-filter(cleaned_classifications, subject_ids %in% correct_subjects)
filesToCheck<-filter(cleaned_classifications, subject_ids %in% subjects_to_filter)
```

write correct file
```{r}
clean_output<-here("Output", "PostZooniverse", "Consensus", "Correct")

correctFile<-paste0(file_name_parts[1], "_", file_name_parts[2],"_", file_name_parts[3], "_correct.csv")

write.csv(correct_classifications, paste0(clean_output,"/", correctFile), row.names = FALSE)
#row.names=FALSE prevents R from adding X column in output df
```

write file that needs checking
```{r}
needs_checking_output<-here("Output", "PostZooniverse", "Consensus", "Needs check")

needToCheckFile<-paste0(file_name_parts[1], "_", file_name_parts[2],"_", file_name_parts[3], "_needToCheck.csv")

write.csv(filesToCheck, paste0(needs_checking_output,"/", needToCheckFile), row.names = FALSE)
#row.names=FALSE prevents R from adding X column in output df
```


And now summarize
```{r}
#these may need adjusting based on what columns are present in the particular file
filesToCheckSummary<-filesToCheck %>% group_by(subject_ids) %>% reframe(
  workflow_name = workflow_name,
  workflow_version = workflow_version,
  choice = choice,
  Img1 = Img1,
  Img2 = Img2,
  Img3 = Img3,
  Forest = Forest1,
  num_class = num_class,
  num_diff_species = num_diff_species
)

```
Now loop through filesToCheck to grab just the first instance of each subject id and associated variables.  

Make data frame to put results in 
```{r}
Subjects<-unique(filesToCheckSummary$subject_ids)

Checklist<-data.frame(subject_ids = integer(length(Subjects)),
                      Img1 = character(length(Subjects)), 
                      Img2 = character(length(Subjects)), 
                      Img3 = character(length(Subjects)),
                      Forest = character(length(Subjects)),
                      num_class = integer(length(Subjects)),
                      num_diff_species = integer(length(Subjects)))
```
Now loop
```{r}
for(i in 1:length(Subjects)){
  mySample<-filter(filesToCheckSummary, subject_ids == Subjects[i])
  Checklist$subject_ids[i]<-Subjects[i]
  Checklist$Img1[i]<-unique(mySample$Img1)
  Checklist$Img2[i]<-unique(mySample$Img2)
  Checklist$Img3[i]<-unique(mySample$Img3)
  Checklist$Forest[i]<-unique(mySample$Forest)
  Checklist$num_class[i]<-unique(mySample$num_class)
  Checklist$num_diff_species[i]<-unique(mySample$num_diff_species)

}
```

Add column for final species ID

```{r}
Checklist$Final_choice <- NA
```

Now save

```{r}

checklist_output<-here("Output", "PostZooniverse", "Consensus", "Needs check", "Checklists")

ChecklistFile<-paste0(file_name_parts[1], "_", file_name_parts[2],"_", file_name_parts[3], "_checklist.csv")

write.csv(Checklist, paste0(checklist_output,"/", ChecklistFile),row.names = FALSE)
#row.names=FALSE prevents R from adding X column in output df

```

Now we need to assign the correct species for the files that we are not checking:
Extract from cleaned classifications all of the subject ids that are NOT part of the list that needs to be checked:

```{r}
Subjects_to_keep<-species_votes %>% filter(propvote >= 0.8)
Subjects_to_keep <-Subjects_to_keep$subject_ids


NeedAssignment<-cleaned_classifications %>% filter(subject_ids %in% Subjects_to_keep)
```
Now create data frame to accept final assignments

```{r}
Assigned<-data.frame(subject_ids = integer(length(Subjects_to_keep)),
                Img1 = character(length(Subjects_to_keep)), 
                Img2 = character(length(Subjects_to_keep)),
                Img3 = character(length(Subjects_to_keep)),
                Img4 = character(length(Subjects_to_keep)), 
                Img5 = character(length(Subjects_to_keep)),
                Img6 = character(length(Subjects_to_keep)),
                Forest = character(length(Subjects_to_keep)),
                num_class = integer(length(Subjects_to_keep)),
                num_diff_species = integer(length(Subjects_to_keep)),
                Final_choice = character (length(Subjects_to_keep)))
```
Now loop
```{r}
NeedAssignment$choice<-as.character(NeedAssignment$choice)
for(i in 1:length(Subjects_to_keep)){
  mySample<-filter(NeedAssignment, subject_ids == Subjects_to_keep[i])
  Assigned$subject_ids[i]<-Subjects_to_keep[i]
  Assigned$Img1[i]<-unique(mySample$Img1)
  Assigned$Img2[i]<-unique(mySample$Img2)
  Assigned$Img3[i]<-unique(mySample$Img3)
  Assigned$Img4[i]<-unique(mySample$Img4)
  Assigned$Img5[i]<-unique(mySample$Img5)
  Assigned$Img6[i]<-unique(mySample$Img6)
  Assigned$Forest[i]<-unique(mySample$Forest1)
  Assigned$num_class[i]<-unique(mySample$num_class)
  Assigned$num_diff_species[i]<-unique(mySample$num_diff_species)
  Assigned$Final_choice[i]<-unique(mySample$choice[1])
 
}
```
Now save

```{r}
assigned_output<-here("Output", "PostZooniverse", "Consensus", "Correct", "Assigned")

assignedFile<-paste0(file_name_parts[1], "_", file_name_parts[2],"_", file_name_parts[3], "_assigned.csv")

write.csv(Assigned, paste0(assigned_output,"/", assignedFile),row.names = FALSE)
#row.names=FALSE prevents R from adding X column in output df

```

## Do final check to make sure that all subject ids are covered
The Checklist output file and the Assigned output files together should include all of the subject_ids that are present in the original classification file.

Let's to a test to make sure.

First, extract the list of subject_ids in the flattened file, the assigned file and the checklist file:
```{r}
All_subjects <-unique(Flat$subject_ids)
Assigned_subjects<-unique(Assigned$subject_ids)
Checklist_subjects <- unique(Checklist$subject_ids)
```
Next, append the subjects from Assigned and Checklist together
```{r}
Account_for <- c(Assigned_subjects, Checklist_subjects)
```

Then find the elements of all subjects that are not present in account for
```{r}
StillNeedToCheck <- setdiff(All_subjects, Account_for)
```

And look at those subjects in the main data frame

```{r}
WhichOnes <- cleaned_classifications %>% filter(subject_ids %in% StillNeedToCheck)
```
Now join to species_votes_to_check to see what the prop vote was 

```{r}
LastOnes <-left_join(WhichOnes, species_votes)

```
Now simplify that file

Create data frame 

```{r}
LastToCheck<-data.frame(subject_ids = integer(length(LastOnes)),
                Img1 = character(length(LastOnes)), 
                Img2 = character(length(LastOnes)),
                Img3 = character(length(LastOnes)),
                Forest = character(length(LastOnes)),
                num_class = integer(length(LastOnes)),
                num_diff_species = integer(length(LastOnes)),
                Final_choice = character(length(LastOnes)))
```
Now loop
```{r}
ones_to_check <-unique(LastOnes$subject_ids)
for(i in 1:length(LastOnes)){
  mySample<-filter(LastOnes, subject_ids == ones_to_check[i])
  LastToCheck$subject_ids[i]<-ones_to_check[i]
  LastToCheck$Img1[i]<-unique(mySample$Img1)
  LastToCheck$Img2[i]<-unique(mySample$Img2)
  LastToCheck$Img3[i]<-unique(mySample$Img3)
  LastToCheck$Forest[i]<-unique(mySample$Forest1)
  LastToCheck$num_class[i]<-unique(mySample$num_class)
  LastToCheck$num_diff_species[i]<-unique(mySample$num_diff_species)
  LastToCheck$Final_choice[i]<-NA
 
}
```


Now write that file to run check


```{r}
last_ones_to_check<-here("Output", "PostZooniverse", "Consensus", "Needs check", "Last Ones")

LastOnesFile<-paste0(file_name_parts[1], "_", file_name_parts[2],"_", file_name_parts[3], "_lastones.csv")

write.csv(LastOnes, paste0(last_ones_to_check,"/", LastOnesFile),row.names = FALSE)
#row.names=FALSE prevents R from adding X column in output df
```





