---
title: "1.Load_data_from_googledrive"
format: html
editor: visual
---

## Overview

Our classification files from the Zooniverse get larger and larger (currently \> 700 MB) which clogs both pushing to GitHub (due to large file size issue) and drive space.

This code allows you to point to a classification .csv file on the Zooniverse, subset it to the desired workflows, and do additional filtering so that you can write a smaller .csv file to use for analysis (or to keep in memory while you use the project but not write).

Using the scripts here requires the `googledrive` package.

## Set things up

```{r}
#| output: false
rm(list = ls())
library(tidyverse)
library(googledrive)
library(here)
```

1. Provide url to file on google drive (you must have at least "viewer" permission to see the file).

```{r}
file_id <- "https://drive.google.com/file/d/1LRyTcWR1xTyPr6MkvG8crqNNgL68CeAj/view?usp=drive_link"
#need to pass the shared url for the correct csv file here
```

2. Now get some info about the file. Requires `googledrive`

```{r}
my_file <- drive_get(file_id) #pulls info from google drive for the file
my_file$name 
#prints file name - check to make sure you have pulled correct file. But first will ask you to authenticate with google drive.
```
3. Now read the file
```{r}
df <-read_csv(drive_read_string(my_file))
#may be very slow!
```

## Subset to correct workflow and version

The Zooniverse uses different "workflows," each of which can have different versions (as they are tweaked), for classfying images. 

We start by identifying the workflows present in our classification file

```{r}
unique(df$workflow_id)
unique(df$workflow_name)
```

The workflow we need to use is called "Species identification"

Let's filter just to that

```{r}
df <- df |> filter(workflow_name == "Species identification")
```

Now we need to see how many versions there are of that worflow

```{r}
unique(df$workflow_version)
```

Generally, we'll want ids from the most current workflow version, in this case 32.13 so we filter to that.

```{r}
df <- df |> filter(workflow_version == 32.13)
```

This reduces us to a much smaller subset of the data.

Next, refer to the file called "3.Flatten_Species_IDs" to see how we get the classification data from the file.

It assumes that you still have `df` in your workspace.
