---
title: "Flattening our Zooniverse files"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

**THIS FILE IS FOR PARSING THE JSON DATA FROM THE "SPECIES IDENTIFICATION" WORKFLOWS. USE THE FILE "FLATTEN_IS_THERE_ANIMAL_ZOONIVERSE.RMD" TO PARSE ANIMAL PRESENCE/ABSENCE DATA.**

## Overview
Output from classifications on the [Zooniverse](https://zooniverse.com) come encoded in *JSON format*.  While most of the columns are straightforward to interpret, two critical columns are not.  The first is the **annotation column**.  Within it are the species identification and the answers to all of the questions for each event. Likewise, the **subject_data** column contains all of the information from the manifest, such as the forest type, photo batch, camera number, and forest name.  It would be good to be able to pull all of this information. We especially need it if we plan to remove e.g. images taken on malfunctioning cameras or otherwise suspect images.
 
Processing the data requires five steps:
1. Selecting the correct classification file.  That file is obtained from the `1.Parse_workflows.Rmd` file in the previous step. **For this code to run, we are assuming you have a single classification file containing only one workflow and single workflow version.**

2. Parsing the annotations column. The annotations column is the column in the zooniverse classification file that includes all of the answers to the questions in the zooniverse workflow, including what the species is and what the animals are doing.

3. Parsing the subject_data column. The subject_data column is the column from the zooniverse classification file that includes information about each subject set from the manifest file.

4. Join the annotation & subject_data together into a data frame for further data analysis.

5. Write the output to a folder called "Flattened" in the PostZooniverse Output folder.


### Turn on packages and bring in the data and r scripts
We need to load desired packages and also source custom functions  so that they will run. The tidyjson package may have to be loaded from this [github](https://github.com/sailthru/tidyjson) site. Look [here](https://cran.r-project.org/web/packages/githubinstall/vignettes/githubinstall.html) for some hints as to how to do that.


If you are starting from scratch and want to clear R's environment:
```{r, message = FALSE, echo = FALSE}
rm(list = ls())
library(tidyverse)
library(here)
```

Otherwise, start here (e.g. if coming from `1.Load_data_from_googledrive`):

```{r, message=FALSE, echo=FALSE}
library(tidyjson)#may get error; if so, load from github
library(jsonlite)
library(magrittr) #allows piping beyond tidyverse
library(stringr)
library(tidyr)
library(rstudioapi)


myFunctions<-list.files(here("Code", "Functions"), pattern = "R") 
#points to files that need to be sourced

invisible(sapply(here("Code", "Functions",myFunctions), source)) #sources all of the functions in the Functions folder. 
#invisible function sources the functions without printing value ? and visible FALSE in the console for each function that is sourced.

```

###Step 1. Choose the classification file to work with.
**If you already reduced your data to a single classification and workflow, you can skip this step and move down to Step 2 below.**

We have multiple classification files as a result of the `1.Parse_workflows.Rmd` file operation. 

Here we point to the folder containing those files and obtain information about the file. 

```{r}
DF<-choose_classification_file()
data<-DF[[1]]
filename<-DF[[2]]
```

### Step 2. Obtain metadata about the file

#### Method A. Assumes you are running this script from the top and didn't already have a dataframe in memory.

Let's gather some information about this data file and create a data frame in which to gather it.

```{r}
File_metadata <- data.frame(Parameter = character(),
                            Value = character()) #intialize empty dataframe

File_metadata[1,] <- c("File name", filename)
File_metadata[2,] <- c("Workflow ID", unique(data$workflow_id))
File_metadata[3,] <-c("Workflow name", unique(data$workflow_name))
File_metadata[4,] <-c("Workflow version", unique(data$workflow_version))
File_metadata[5,] <-c("Num unflattened rows", length(data$classification_id))

```

#### Method B. Assumes you have a single dataframe read in via e.g. the `googledrive` package.

```{r}
File_metadata <- data.frame(Parameter = character(),
                            Value = character()) #intialize empty dataframe

File_metadata[1,] <- c("File name", my_file$name)
File_metadata[2,] <- c("Workflow ID", unique(df$workflow_id))
File_metadata[3,] <-c("Workflow name", unique(df$workflow_name))
File_metadata[4,] <-c("Workflow version", unique(df$workflow_version))
File_metadata[5,] <-c("Num unflattened rows", length(df$classification_id))
File_metadata[6,] <- c("Num unique classifications", length(unique(df$classification_id)))
File_metadata[7,] <- c("Num unique subjects", length(unique(df$subject_ids)))
```


### Step 3. Parse the annotations column
We now flatten the Annotations column by calling the  `flatten_annotations` function. 

```{r}
flat_file<-flatten_annotations(df) #this is slow; could we write faster code?  A problem for another day.
```
Why does the flattened file have fewer rows than the initial dataframe?

Let's see which `subject_ids` in the main data frame are missing from the flattened file:

```{r}
not_in_flat <- setdiff(df$classification_id, flat_file$classification_id)
```
You can see that the difference are rows of data for which a classification was not completed.

So, flat_file has all of the data for which a classification was actually made.

### Step 4. Parse the subjects column

Now we can parse the data.

If you are working with a recent version of the species identification workflow, use the `flatten_subjects_wflow_32_13.R` function.
```{r}
subjects<-flatten_subjects_wflow_32_13(df)
```


If you are using data from an older workflow, there may be some wonky issues when you flatten subjects -- use the basic `flatten_subjects` function if that is the case.

I've updated flatten_subjects so that it reports a lot of columns, parsing all of the subject columns regardless of the format of the manifest that was uploaded to the zooniverse. So now we can sort by project and isolate data from different manifests that use the same workflow. 

Notice that if flat_file may have fewer rows than subjects it is because for some subset of the data (equal in number to the difference in length b/w the subjects data and the flat_file data) no identification was made, so the annotations column has no information.

###Step 5. Merge the annotations and subjects data
Now that we have flattened both the annotations and subject_data columns, we would like to generate one large data frame with all of the data to export and analyze (using different R scripts!).  To do so, we need to join the two data frames.  Joining will **only work** if you have a column, named identically in both data frames, on which the join will work.

The join itself is pretty easy:
```{r}
Flattened<-left_join(flat_file, subjects, by='classification_id')

# Now move subject_ids column to left most position

Flattened<-Flattened %>% relocate(subject_ids)

#and move the Camera_SD card combination toward the left as well.

Flattened <- Flattened |>
  relocate(CamSD, .after = classification_id)
```

## Save the result

Now Save our result! This will go in the flattened folder within the PostZooniverse outputs.

### Method A. Assumes you are running this script from the top and didn't already have a dataframe in memory.
```{r}
output<-here("Output", "PostZooniverse", "Flattened", "Species_ID")

myFile<-str_remove(filename, ".csv") #drop the .csv temporarily

write.csv(Flattened,  paste0(output,"/",myFile,"_flat_", Sys.Date(), ".csv"))

#and write the metadata
write.csv(File_metadata, paste0(output,"/",myFile,"_fMeta_", Sys.Date(), ".csv"))

```

### Method B. Write the file locally and then push to a google drive.

#### Step 1. Save the .csv file
```{r}
write_csv(Flattened, here("Output/PostZooniverse/Flattened/Species_ID/Flattened-wf16995_32_13-11-Mar-2026.csv"))
```

#### Step 2. Push to google drive
Will write this code later

#### Step 3. Remove file from local project if using github repo
Because of the "large file size" issue with GitHub, if you have written the file (Step 1 above) to a folder within the repo, you may have problems pushing commits to GitHub if you make any project commits while the large file is part of the project. So you may wish to delete the local version of the file from the project.

